$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2014-04-07 JournalToSubmit2013/modelingFlyingPig.tex
Answers for Martin and Umberto.

$\star$ "Maximum number of users should be greater than 200".
}
\caption{$\pi$-UseCase model for \FlyingPig.\label{fig:piUseCaseModel}}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\textwidth]{figs/ServiceCompositionGeneralCut.png}
{\color{red} \raggedright
$\star$ Change FlyingFish by \FlyingPig.
}
\caption{$\pi$-ServiceComposition model for \FlyingPig.\label{fig:PiServiceCompositionModel}}
\end{figure}

>>>>

$\star$ "Maximum number of users should be greater than 200".
} {\color{blue} [[I did not understand what must be done.]]}
\caption{$\pi$-UseCase model for \FlyingPig.\label{fig:piUseCaseModel}}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\textwidth]{figs/ServiceCompositionGeneralCut}
{\color{red} \raggedright
$\star$ Change FlyingFish by \FlyingPig.
} {\color{blue} [[done.]]}
\caption{$\pi$-ServiceComposition model for \FlyingPig.\label{fig:PiServiceCompositionModel}}
\end{figure}

### 2014-04-07 JournalToSubmit2013/transformationRules.tex
Answers for Martin and Umberto.

As defined by the $\pi$-ServiceProcess  meta-model (Figure~\ref{fig:CIM:serviceprocessmetamodel}) an {\sc Activity Service} consists of a composition of  entities of type {\sc Action}. 
Thus, every {\sf $\pi$-use case} is transformed into an {\sf Action} of the target $\pi$-Service Process model.  
Every {\sf Extend} relationship identified in a $\pi$-Use Case model is transformed into a  {\sf Fork node}\footnote{Notice that Fork nodes are used both to represent alternative and parallel execution. {\color{red} \sc Placido, please check this!}}.
If the {\sf Extend} relationship concerns just one {\sf use case}, it is transformed into an {\sf Action} inside one flow of the fork node. 
Otherwise, several  {\sf use cases} are transformed into different {\sf Actions} that belong to different flows departing from the fork node.

@@

Placido: Is this correct? Why it is here? We consider the $\pi$-use case in the previous section!
}

@@

An {\sf Exceptional behaviour} entity is transformed into an {\sf Action} in a $\pi$-Service Composition model and every {\sf Non-functional attribute} associated to an element ({\sf Contract} or {\sf Non-functional requirement}) in a $\pi$-service process model becomes a {\sf Non-functional attribute} associated to the corresponding element {\sf A-Policy} in a $\pi$-service composition model.\footnote{\color{red} Not found in the metamodel! --Martin and Umberto}
Finally, {\sf Actions} are grouped by  a {\sf Business collaborator}.\footnote{\color{red} How? Where in the metamodel? --Martin and Umberto}

@@

The "securityLoginPolicy" consists of a set of {\sf Rules} that were transformed from the {\sf Assertions} in $\pi$-service process model. 
The information about Facebook and Spotify (both of them {\sf Business Collaborators}) come from entity of type {\sc Package} in the $\pi$-use case model.\footnote{\color{red} Como assim? - Placido, explain this, please! Why have we bypassed one level here?}
\end{example}

@@

Single actions (represented by {\sf Action} and {\sf Action:name}\footnote{\color{red} Where is the name defined?}) are transformed into individual service {\sf Operations}. 
Complex actions (represented by {\sf ServiceActivity}  and  {\sf ServiceActivity:name}\footnote{\color{red} Where is the name defined?}) are transformed into (named) composite operations that defines a (named) workflow of the application. 
Composition patterns expressed using the operators {\sc\em merge, decision, fork and join} are transformed into their corresponding workflows of the $\pi$-PEWS model.

@@

The A-Policies defined for the entities of a $\pi$-service composition model are transformed into {\sf A-policy}
\footnote{\color{red} Please rename this for CONTRACT on the pi-pews metamodel!} 
entities, named according to the names expressed in the source model. 
The transformation of the rules expressed in a $\pi$-service composition is guided by the event types associated to these rules. 
The variables associated to an A-Policy expressed in a $\pi$-service composition model as {\sf $<$Variable:name, Variable:type$>$} are transformed into entities  {\sf Variable} with attributes {\sf Name} and {\sf Type} directly specified from the elements {\sf Variable:name} and {\sf Variable:type} of a $\pi$-service composition model.

@@

\item 	We  used Acceleo\footnote{http://www.acceleo.org/pages/home/en} for implementing  the code generation plug-in for generating executable code. 
It takes a $\pi$-PEWS model and generates the code to be executed by the {\em A-Policy} based service composition execution environment\footnote{\color{red} Placido: Policy or contract??.}. 
\end{itemize}

>>>>

As defined by the $\pi$-ServiceProcess  meta-model (Figure~\ref{fig:CIM:serviceprocessmetamodel}) an {\sc Activity Service} consists of a composition of  entities of type {\sc Action}. 
Thus, every {\sf $\pi$-use case} is transformed into an {\sf Action} of the target $\pi$-Service Process model.  
Every {\sf Extend} relationship identified in a $\pi$-Use Case model is
transformed into a  {\sf Fork node}\footnote{Notice that Fork nodes are used
both to represent alternative and parallel execution. {\color{red} \sc Placido,
please check this!} {\color{blue} This is consequence of the use cases' extends
and include relationships. To transform an use case into an action, it is
necessary check the dependence relation. The include implies a parallel fork,
and a extends implies in a alternative fork. I included the references for this definition.
} }\cite{valeriaThesis,placidoPhDThesis2012}.
If the {\sf Extend} relationship concerns just one {\sf use case}, it is transformed into an {\sf Action} inside one flow of the fork node. 
Otherwise, several  {\sf use cases} are transformed into different {\sf Actions} that belong to different flows departing from the fork node.

@@

Placido: Is this correct? Why it is here? We consider the $\pi$-use case in the previous section!
}{\color{blue} Yes, it is correct. I defined it in the thesis because in the
previous SOD-M a Business Collaborator comes from 'nothing'. So, as we needed to
refine most concepts from use case, I decided separete the use cases in
packages and then tranform it in a Collaborator. As a use case is transformed
into an action, and all action is from a BC, I related both concepts. There is a
layer gap ($\pi$-ServiceProcess), because in this layer we define only the
actions.

We are defining it here because the concept is necessary for the
transformation. Packages is part of UML in the use case model, and I used it for define the colaborators.}

@@

{\color{blue} An {\sf Exceptional behaviour} entity is transformed into an {\sf
Action} in a $\pi$-Serv\-ice\-Com\-po\-si\-tion model and every {\sf
Non-functional attribute} associated to an {\sf Contract} (it comes from the $\pi$UseCase model) element 
in a $\pi$-ServiceProcess model becomes a {\sf Non-functional Requirement}
associated to the corresponding element {\sf A-Policy} in a
$\pi$-ServiceComposition model}.\footnote{\color{red} Not found in the
metamodel! --Martin and Umberto \color{blue}  ANSWER:  If you analyze the model $\pi$-UseCase, a constraint may be associated with one or more NFA (Non-functional attribute). The constraints
associated with the same use case, are grouped into a contract. Each contract is
related to a actions ($\pi$-ServiceProcess). Thus, the contracts (set of
constraints) have NFAs associated (from piUseCase). The transformation between
the first 2 levels (PIM $\pi$-UseCase to $\pi$-ServiceProcess) give us this.
NFAs associated to an element (Contract) in a $\pi$-ServiceProcess model are
refined (grouped) into a Non-functional requirement associated to the corresponding
element A-Policy in the $\pi$-ServiceComposition model. I have to update the
meta-models (thesis version) because they were outdated.} Finally, {\sf Actions}
are grouped by  a {\sf Business collaborator}.\footnote{\color{red} How? Where
in the metamodel? --Martin and Umberto \color{blue} ANSWER: Figure
\ref{fig:e-scomposition-metamodel} presents the relationship between Action and BusinessCollaborator}

@@

The "securityLoginPolicy" consists of a set of {\sf Rules} that were transformed from the {\sf Assertions} in $\pi$-service process model. 
The information about Facebook and Spotify (both of them {\sf Business
Collaborators}) come from entity of type {\sc Package} in the $\pi$-use case
model.\footnote{\color{red} Como assim? - Placido, explain this, please! Why
have we bypassed one level here? \color{blue} ANSWER: It was explained before.}
\end{example}

@@

Single actions (represented by {\sf Action} and {\sf
Action:name}\footnote{\color{red} Where is the name defined? \color{blue} ANSWER: I updated the metamodel with the name atribute, however
in general, I ommit the atributes for the metamodel presentation, but in the
implementation the atributes are there.}) are transformed into individual service {\sf Operations}.
Complex actions (represented by {\sf ServiceActivity}  and  {\sf
ServiceActivity:name}\footnote{\color{red} Where is the name defined?
\color{blue} ANSWER: I updated the metamodel with the name atribute, however
in general, I ommit the atributes for the metamodel presentation, but in the
implementation the atributes are there.}) are transformed into (named) composite
operations that defines a (named) workflow of the application.
Composition patterns expressed using the operators {\sc\em merge, decision, fork and join} are transformed into their corresponding workflows of the $\pi$-PEWS model.

@@

The A-Policies defined for the entities of a $\pi$-service composition model are transformed into {\sf A-policy}
\footnote{\color{red} Please rename this for CONTRACT on the pi-pews metamodel!
\color{blue} ANSWER: The latest version of the $\pi$-PEWS grammar I defined
policy.} entities, named according to the names expressed in the source model.
The transformation of the rules expressed in a $\pi$-service composition is guided by the event types associated to these rules. 
The variables associated to an A-Policy expressed in a $\pi$-service composition model as {\sf $<$Variable:name, Variable:type$>$} are transformed into entities  {\sf Variable} with attributes {\sf Name} and {\sf Type} directly specified from the elements {\sf Variable:name} and {\sf Variable:type} of a $\pi$-service composition model.

@@

\item 	We  used Acceleo\footnote{http://www.acceleo.org/pages/home/en} for implementing  the code generation plug-in for generating executable code. 
It takes a $\pi$-PEWS model and generates the code to be executed by the {\em
A-Policy} based service composition execution environment\footnote{\color{red}
Placido: Policy or contract??. \color{blue} ANSWER: The latest version of the
$\pi$-PEWS grammar I defined \textbf{policy}. Particularly I prefer \textbf{policy},
because for what we're proposing, one same policy can be applied to different
services and groups a set of contracts.}.
\end{itemize}
